\tool: Programmable, Efficient, and Scalable Secure Two-Party Computation

### 2017-10-26 OaklandSubmission/intro.tex

Since its introduction in 1980s, there has been a long sequence
of work that has transformed \mpc from a mere theoretical concept to
something practically efficient and usable. To understand the
state-of-the-art, let us consider the following example used by many
secure prediction algorithms. Suppose Alice wants to write
a \mpc protocol to securely
compute $w^Tx >b$. Here $w$ and $x$ are two vectors (inputs of the two
parties), $b$ is a scalar, $\cdot^{T}$ is the matrix transpose
operator, and $w^Tx$ denotes the inner product of
$w^{T}$ and $x$.  With the current state-of-the-art, Alice has three
options:

@@

She can program the computation in one of the several programmer
friendly, domain-specific languages (such
as~\cite{lambdaps,wysteria,oblivm,smcl,cbmcgc}) that would
automatically compile it to a \mpc protocol. However, all of these
frameworks use cryptographic backends that take as input
the computation expressed as either a boolean
circuit~(\cite{yao,gmw}) or an arithmetic circuit~(\cite{gentry}).
%
The efficiency of the generated \mpc protocol is thus bounded by the
efficiency of representing the computation in one of
these representations. For instance, multiplication of two
$\ell$-bit integers can be either expressed as a boolean circuit of
size $O(\ell^2)$, or as an arithmetic circuit with 1 multiplication
gate.
%
Hence, for better efficiency, Alice would ideally like to compute
$w^Tx$ using an arithmetic circuit, and the comparison with $b$ using
a boolean circuit.
%
Unfortunately, none of the above frameworks support combinations of
arithmetic and boolean circuits, and using different tools for
different parts of the computation is cumbersome and error-prone.

@@

Alternatively, Alice can use a tool such as ABY (Demmlet et
al.~\cite{aby}) that allows
the computation to be expressed as a combination of arithmetic and
boolean circuits. However, here the programming interface is quite
low-level -- the programmer is required to first manually split the
computation into arithmetic and boolean components, and then write the
circuits for all the componenets manually, including the appropriate
inter-conversion gates between them. Clearly, writing correct
and efficient protocols in such a framework is beyond an average
programmer who does not understand the various trade-offs between
arithmetic and boolean circuits, and even for an expert cryptographer,
writing large computations in such a framework can be tedious and
error-prone (a sentiment also echoed by Demmler et al.~\cite{aby}).

%\item
The third option for Alice is to earn a PhD in cryptography and design
specialized, efficient protocols (similar to
~\cite{shafindss,wu,minionn}) for her tasks.
%\end{tiret}
%In our work, we achieve the best of all the above options.

@@

\subsection{\tool: Overview and Contributions} 
\label{sec:contrib}
%Write high level overview of programming language and tool.. \divya{have some preamble}

In this work, we develop and implement our framework \tool\footnote{\tool, read as ``easy peasy'', stands for Easy 2 Party Computation} for secure
computation. We provide the first compiler where the developer writes a high-level program and the compiler automatically generates an efficient \mpc protocol that uses a
mix of arithmetic and boolean circuits. 
It is compatible with any implementation of \mpc that allows a mix of arithmetic and boolean circuits. 
We focus on semi-honest secure two-party computation and use ABY \cite{aby} as the suitable cryptographic back-end.
We provide a formal type system and operational semantics for our language and prove correctness and security of our compiler.
We evaluate our system on wide range of benchmarks for secure prediction \cite{shafindss,wu,barni,secureml,minionn,bonsai} using linear classifier, decision trees and deep neural networks, matrix factorization \cite{valeriaMatrix} and training of deep neural network \cite{secureml}. Our evaluations show that protocols generated by our compiler match
the performance or outperform hand-crafted protocols in most cases (see
\sectionref{eval}). \TODO{Add some text to increase enthusiasm about results.} Moreover, unlike previous systems such as \cite{wysteria,aby,cbmcgc}, our system can scale to arbitrarily large computations using our technique of {\em secure code partitioning} (see below). \divya{I think we can skip Yao pipelining here, right?}
Hence, we show that our framework achieves all four properties of programmability, generality, performance, and scalability (see below for details).

Prior to our work, there have been systems that provide only a subset of the above properties. For instance, using the framework of CBMC-GC \cite{cbmcgc} allows  functionalities to be expressed in high-level language C but it gives poor performance due to its usage of boolean circuits as underlying representation. Similarly, ABY \cite{aby} framework can potentially give good performance (due to mix of boolean and arithmetic circuits) but is not programmable. Yao's garbled circuits \cite{yao} work with boolean circuits and are the most widely used 2PC protocols but they  do not scale to large functions due to large memory requirement.
Huang et al. \cite{yao-pipe} gave the technique of pipelining Yao's garbled circuits that enabled scaling to large functions but still gives poor performance due to its use of boolean circuit representation. 
To summarize, none of the prior frameworks have been used to evaluate the benchmarks such as secure prediction using linear regression, decision trees, deep neural networks, etc. In fact, it was widely believed that generic 2PC would not scale to the secure prediction and hence, various protocols were hand-crafted for this task \cite{shafindss,wu,secureml,minionn}. Our framework contradicts this belief (see below).

In this work, we solve the design problem to provide a user friendly language that is expressive and can be compiled to efficient secure 2PC protocols.
Programmability in \tool and performance of generated protocols has enabled us to write large programs whose circuits  exhaust the memory of the system (~32 GB). To combat this issue, we give a technique called secure code partitioning (see below) that allows scaling of \tool to arbitrary size computations.

Below, we describe the salient features of \tool in more detail. \\

\noindent\textbf{Programmability.} \tool is a programmer-friendly framework to
express the function to be computed and uses a mix of arithmetic and
boolean \mpc back-ends. We illustrate the above example of $w^Tx
>b$ in our language in \figureref{ex-sml}. 
As is clear from the code, the programmer can remain completely oblivious of the cryptographic back-end and underlying circuit representation. 
Also,  the programmer is not required to write keywords such as {\em secret}
or {\em public} (that are required by languages like Wysteria \cite{wysteria}). 
 In contrast, \figureref{ex-aby} illustrates
the same example using the framework in ABY \cite{aby}. Here, as
already mentioned, the programmer needs to create shares and write the low level gates
like $\mathtt{MUL, ADD, CONS, GT}$, etc \TODO{make these fonts consistent}. The programmer is also required
to use the appropriate share conversion gates such as $\mathtt{A2Y}$ that
converts arithmetic shares to Yao-based boolean shares to be used in comparison after the inner-product computation.
Moreover, it is easy to
maintain code in \tool and a small change to the functionality requires
only a small change in the code. In contrast, working with ABY, even
for a small change a developer might have to re-work the whole circuit
for efficiency reasons. We elaborate on these examples  in
\sectionref{ex}. \\

\noindent\textbf{Generality and Performance.} Our compiler is aware of cryptographic costs of different representations and corresponding inter-conversions.  Based on these costs, our compiler, given a program in \tool, automatically picks arithmetic and boolean representations for different sub-parts of the program and inserts the required inter-conversions as well. 
We demonstrate the generality of our framework by writing a wide range of functionalities in \tool such as linear regression, deep neural networks, decision trees and matrix factorization.
Our evaluations show that automatically generated protocols using \tool are competitive with tailor-made protocols in \cite{shafindss,wu,barni,minionn,secureml,valeriaMatrix} (see \sectionref{eval}). These features allow us to evaluate our system state-of-the-art ML prediction algorithms, e.g., \sectionref{eval} describes evaluation on ML models from a recent ICML paper \cite{bonsai} and Tensorflow \cite{tensorflow}. \\

\noindent\textbf{Performance Baseline.} Most works cite garbled circuit based two-party computation as the most practical approach so far and use it as the baseline for comparison of performance. But, it has been shown multiple times that the garbled circuits based approach does not scale for realistic examples. Even for small benchmarks (e.g., Na\"{i}ve Bayes) the size of garbled circuit can be huge. Hence, comparisons are often done on artificial data sizes and garbled circuits are shown to be 500x slower than hand-crafted protocols \cite{shafindss}. This is cited as the motivation for need for specialized protocols for different ML algorithms in secure prediction such as linear regression \cite{shafindss}, decision trees \cite{wu}, deep neural networks \cite{minionn} and matrix factorization \cite{valeriaMatrix}. Our work is based on generic \mpc and we show competitive performance with the specialized protocols contrary to the popular belief. Hence, we believe \tool based implementations can serve as the performance baseline for future works on specialized protocols. \\

@@

\noindent\textbf{Static Guarantees.} We give a formal {\em type system} for our language and prove useful properties. We prove that if a program type checks in our language then it would run correctly, e.g., there would not be any array index out of bounds error. The programmer writes a single code describing the inputs from each party and the function to be computed. Our compiler automatically generates separate executables for the server and the client that can be executed in the distributed setting. We prove the {\em correctness} for this compilation, i.e., the outputs obtained in the distributed execution are identical to the stand-alone execution by a trusted third party. We formally reduce the {\em security} of our system to the security of the cryptographic back-end. For details on formal language, correctness and security theorems see \sectionref{ld}. \\

\noindent\textbf{Secure Code Partitioning (or Partitioning in short).} Another issue in adoption of \mpc is difficulty in scaling to large programs. 
%In general, \mpc implementations struggle in scaling to large programssize of memory on a system limits the largest compute inside the SMC and hence,. 
The reason being most \mpc implementations use a circuit-like representation as an intermediate representation. Hence, the largest compute that can be done securely is upper-bounded by the largest circuit that can fit in memory on a machine\footnote{Using swap and disk space is feasible but it causes huge slowdown.}.
To address this issue, we develop a technique called partitioning.  At a high level, we break the original program into a sequence of small sub-programs and mask the intermediate outputs using a secure secret sharing between the two parties. We appropriately append the share-reconstruction function to the beginning of the subsequent sub-program. Now, we execute each of these sub-programs sequentially in our framework. This allows us to run arbitrarily large programs. We provide details of partitioning and prove the correctness and security of this transformation in \sectionref{pipe}. \\

>>>>

Since its introduction in 1980s, there has been a long sequence
of work that has transformed \mpc from a mere theoretical concept to
something practically efficient and usable. To understand the
state-of-the-art, let us consider the following example used by many
secure prediction algorithms. Suppose Alice wants to write
a \mpc protocol to securely
compute $w^Tx >b$. Here $w$ and $x$ are two vectors (inputs of the two
parties), $b$ is a scalar, $\cdot^{T}$ is the matrix transpose
operator, and $w^Tx$ denotes the inner product of
$w^{T}$ and $x$.  With the current state-of-the-art, Alice has following
options:

@@

She can program the computation in one of the several programmer
friendly, domain-specific languages (such
as Fairplay~\cite{fairplay}, Wysteria~\cite{wysteria},
ObliVM~\cite{oblivm}, CBMC-GC~\cite{cbmcgc}, SMCL~\cite{smcl},
Sharemind~\cite{sharemind}, \cite{lambdaps} etc.) that would
automatically compile it to a \mpc protocol. However, all of these
frameworks use cryptographic backends that take as input
the computation expressed as either a boolean
circuit~(\cite{yao,gmw}) or an arithmetic circuit~(\cite{gentry}).
%
The efficiency of the generated \mpc protocol is thus bounded by the
efficiency of representing the computation in one of
these representations. For instance, multiplication of two
$\ell$-bit integers can be either expressed as a boolean circuit of
size $O(\ell^2)$, or as an arithmetic circuit with 1 multiplication
gate.
%
Hence, for better efficiency, Alice would ideally like to compute
$w^Tx$ using an arithmetic circuit, and the comparison with $b$ using
a boolean circuit.
%
Unfortunately, none of the above frameworks support combinations of
arithmetic and boolean circuits, and using different tools for
different parts of the computation is cumbersome and error-prone.

@@

Alternatively, Alice can use a tool such as ABY (Demmlet et
al.~\cite{aby}) that allows
the computation to be expressed as a combination of arithmetic and
boolean circuits. However, here the programming interface is quite
low-level -- the programmer is required to first manually split the
computation into arithmetic and boolean components, and then write the
circuits for all the componenets manually, including the appropriate
inter-conversion gates between them. Clearly, writing correct
and efficient protocols in such a framework is beyond an average
programmer who does not understand the various trade-offs between
arithmetic and boolean circuits, and even for an expert cryptographer,
writing large computations in such a framework can be tedious and
error-prone (a sentiment echoed by Demmler et al.~\cite{aby} themselves).

%\item
A third option for Alice is to earn a PhD in cryptography, and design
and implement specialized, efficient \mpc protocols (similar
to~\cite{shafindss,wu,minionn}) for her tasks.
%\end{tiret}
%In our work, we achieve the best of all the above options.

@@

\subsubsection*{\tool}
\label{sec:contrib}
%Write high level overview of programming language and tool.. \divya{have some preamble}

This paper presents \tool\footnote{Read as ``easy peasy'',
stands for Easy 2 Party Computation}, a new \mpc framework.
\tool is the first
``cryptographic-cost aware'' framework that generates efficient and
scalable \mpc protocols, comprising combinations of arithmetic and
boolean circuits, from high-level programs devoid of any
cryptography. Below we describe the salient features of \tool.

\subsubsection*{Programmability} \tool source programs are ideal
functionalities that describe ``what'' computation needs to be done,
rather than ``how'' to do it. In particular, the programmer writes the
high-level computation without thinking about the underlying
cryptographic details. For example, an \tool source
program for $w^Tx >b$ is as follows:

\vspace{0.2cm}
\begin{lstlisting}[language=C]
uint w[30] = input1(); uint b = input1();
uint x[30] = input2();
uint acc = 0;
for i in [0:30] { acc =  acc + (w[i] * x[i]); }
output((acc > b) ? 1 : 0)
\end{lstlisting}

\vspace{0.2cm}
We note that the program is quite similar to what a programmer might
write in a common programming language. The simplicity of the
language comes with its usual benefits: it is easily accessible to the
developers, there are fewer avenues for making mistakes, and it is
easy to maintain and modify the programs. Needless to say, frameworks
that expose low-level circuit APIs to the programmer do not enjoy
these benefits.

\subsubsection*{Cryptogaphic-cost awareness} \mpc compiler is aware
of the cryptographic costs of arithmetic and boolean representations
of the source language operators, and their corresponding
inter-conversion gates. Based on these costs, the compiler, given a
source program, automatically picks arithmetic or boolean
representations for different parts of the program, and generates the
corresponding circuits along with the required inter-conversion
gates. The outcome is an effcient \mpc protocol using a combination of
arithmetic and boolean circuits, while the programmer remains
oblivious of all the low-level cryptographic details. Our evaluation
shows that the performance of the
protocols automatically generated by the \tool compiler is comparable
to or better than the custom, specialized protocols developed in the
previous work. Indeed, \tool is the first such cryptographic-cost
aware compiler.

\subsubsection*{Scalability (secure code partitioning)} \mpc tools
often do not scale to large programs. The reason is that
most \mpc implementations use a circuit-like representation as an
intermediate language. Hence, the largest compute that can be done
securely is upper-bounded by the largest circuit that can fit in the
memory on a machine\footnote{Using swap and
disk space is feasible but it causes huge slowdown.}. This is a
show-stopper for applications like secure machine learning, secure
prediction, etc. that operate on large amounts of data. \aseem{Can we
cite something here to support this claim?}
\tool addresses the scalability concern using a novel technique that
we call secure code partitioning (or partitioning in short). At
a high level, we break the original program into a sequence of small
sub-programs, which are then sequentially processed by our
framework, while appropriately threading the intermediate outputs
along. While this
addresses the scalability concern (the circuit
sizes of the sub-programs are now small enough to fit in the memory),
we still have to address
the security risk of revealing the intermediate outputs. To address
this, \tool (a) masks the intermediate outputs using a secure secret
sharing between the two parties, and (b) appends the
share-reconstruction function to the beginning of the subsequent
sub-program. As we show in our evaluation, partitioning allows us to
program large applications in \tool.

%% We provide details of partitioning and prove the
%% correctness and security of this transformation
%% in \sectionref{pipe}

\subsubsection*{Formal guarantees} We prove a formal correctness and
security theorem for our compiler. The correctness theorem relates the
``ideal'' semantics (the trusted-third party semantics) of a source
program and the ``protocol'' semantics (the distributed \mpc
semantics that relies on circuit evaluation) of the corresponding
compiled program. Both these semantics are instrumented
with \emph{observations} that record the outputs. We prove that a
well-typed source program terminates successfully in the source
semantics. In addition, we prove that for a well-typed source program,
its observations in the source semantics are same as the observations
of its corresponding compiled program in the protocol
semantics. {\aseem{Had a hard time writing this last sentence,
suggestions welcome.}}
For the security theorem, we formally reduce the security of our
compiler to the security of the back-end cryptographic
protocol.

%% an intermdiate language (that is the output of our
%% compiler), and a circuit language for arithmetic and boolean
%% circuits.

%% We give a formal {\em type
%% system} for our language and prove useful properties. We prove that if
%% a program type checks in our language then it would run correctly,
%% e.g., there would not be any array index out of bounds error. The
%% programmer writes a single code describing the inputs from each party
%% and the function to be computed. Our compiler automatically generates
%% separate executables for the server and the client that can be
%% executed in the distributed setting. We prove the {\em correctness}
%% for this compilation, i.e., the outputs obtained in the distributed
%% execution are identical to the stand-alone execution by a trusted
%% third party. For details on formal
%% language, correctness and security theorems see \sectionref{ld}.

%% We demonstrate the generality of our framework by writing a wide range
%% of functionalities in \tool such as linear regression, deep neural
%% networks, decision trees and matrix factorization. 
%% Our evaluations show that automatically generated protocols
%% using \tool are competitive with tailor-made protocols
%% in \cite{shafindss,wu,barni,minionn,secureml,valeriaMatrix}
%% (see \sectionref{eval}). These features allow us to evaluate our
%% system state-of-the-art ML prediction algorithms,
%% e.g., \sectionref{eval} describes evaluation on ML models from a
%% recent ICML paper \cite{bonsai} and Tensorflow \cite{tensorflow}. \\

%% \tool is a programmer-friendly
%% framework to
%% express the function to be computed and uses a mix of arithmetic and
%% boolean \mpc back-ends. We illustrate the above example of $w^Tx
%% >b$ in our language in \figureref{ex-sml}. 
%% As is clear from the code, the programmer can remain completely
%% oblivious of the cryptographic back-end and underlying circuit
%% representation.
%% Also,  the programmer is not required to write keywords such as {\em
%% secret}
%% or {\em public} (that are required by languages like
%% Wysteria \cite{wysteria}).
%%  In contrast, \figureref{ex-aby} illustrates
%% the same example using the framework in ABY \cite{aby}. Here, as
%% already mentioned, the programmer needs to create shares and write the
%% low level gates
%% like $\mathtt{MUL, ADD, CONS, GT}$, etc \TODO{make these fonts
%% consistent}. The programmer is also required
%% to use the appropriate share conversion gates such as $\mathtt{A2Y}$ that
%% converts arithmetic shares to Yao-based boolean shares to be used in
%% comparison after the inner-product computation.
%% Moreover, it is easy to
%% maintain code in \tool and a small change to the functionality requires
%% only a small change in the code. In contrast, working with ABY, even
%% for a small change a developer might have to re-work the whole circuit
%% for efficiency reasons. We elaborate on these examples  in
%% \sectionref{ex}. \\

{\aseem{Not sure about this baseline paragraph. Did not touch it.}} 
\subsubsection*{Performance Baseline} Most works cite garbled
circuit based two-party computation as the most practical approach so
far and use it as the baseline for comparison of performance. But, it
has been shown multiple times that the garbled circuits based approach
does not scale for realistic examples. Even for small benchmarks
(e.g., Na\"{i}ve Bayes) the size of garbled circuit can be
huge. Hence, comparisons are often done on artificial data sizes and
garbled circuits are shown to be 500x slower than hand-crafted
protocols \cite{shafindss}. This is cited as the motivation for need
for specialized protocols for different ML algorithms in secure
prediction such as linear regression \cite{shafindss}, decision
trees \cite{wu}, deep neural networks \cite{minionn} and matrix
factorization \cite{valeriaMatrix}. Our work is based on generic \mpc
and we show competitive performance with the specialized protocols
contrary to the popular belief. Hence, we believe \tool based
implementations can serve as the performance baseline for future works
on specialized protocols.

@@

%% We have implemented \tool using ABY~\cite{aby} as the crypto backend.
%% We evaluate our system on wide range of benchmarks for secure
%% prediction \cite{shafindss,wu,barni,secureml,minionn,bonsai} using
%% linear classifier, decision trees and deep neural networks, matrix
%% factorization \cite{valeriaMatrix} and training of deep neural
%% network \cite{secureml}.

%% for secure
%% computation. We provide the first compiler where the developer writes
%% a high-level program and the compiler automatically generates an
%% efficient \mpc protocol that uses a
%% mix of arithmetic and boolean circuits. 
%% It is compatible with any implementation of \mpc that allows a mix of
%% arithmetic and boolean circuits.
%% We focus on semi-honest secure two-party computation and use
%% ABY \cite{aby} as the suitable cryptographic back-end.
%% We provide a formal type system and operational semantics for our
%% language and prove correctness and security of our compiler.
%% Our evaluations show that protocols generated
%% by our compiler match
%% the performance or outperform hand-crafted protocols in most cases (see
%% \sectionref{eval}). \TODO{Add some text to increase enthusiasm about
%% results.} Moreover, unlike previous systems such
%% as \cite{wysteria,aby,cbmcgc}, our system can scale to arbitrarily
%% large computations using our technique of {\em secure code
%% partitioning} (see below). \divya{I think we can skip Yao pipelining
%% here, right?}
%% Hence, we show that our framework achieves all four properties of
%% programmability, generality, performance, and scalability (see below
%% for details).

%% Prior to our work, there have been systems that provide only a subset of the above properties. For instance, using the framework of CBMC-GC \cite{cbmcgc} allows  functionalities to be expressed in high-level language C but it gives poor performance due to its usage of boolean circuits as underlying representation. Similarly, ABY \cite{aby} framework can potentially give good performance (due to mix of boolean and arithmetic circuits) but is not programmable. Yao's garbled circuits \cite{yao} work with boolean circuits and are the most widely used 2PC protocols but they  do not scale to large functions due to large memory requirement.
%% Huang et al. \cite{yao-pipe} gave the technique of pipelining Yao's garbled circuits that enabled scaling to large functions but still gives poor performance due to its use of boolean circuit representation. 
%% To summarize, none of the prior frameworks have been used to evaluate the benchmarks such as secure prediction using linear regression, decision trees, deep neural networks, etc. In fact, it was widely believed that generic 2PC would not scale to the secure prediction and hence, various protocols were hand-crafted for this task \cite{shafindss,wu,secureml,minionn}. Our framework contradicts this belief (see below).

%% In this work, we solve the design problem to provide a user friendly language that is expressive and can be compiled to efficient secure 2PC protocols.
%% Programmability in \tool and performance of generated protocols has enabled us to write large programs whose circuits  exhaust the memory of the system (~32 GB). To combat this issue, we give a technique called secure code partitioning (see below) that allows scaling of \tool to arbitrary size computations.

%% Below, we describe the salient features of \tool in more detail. \\

\aseem{Checkpointing here. Two more paragraphs to come, one for
summary of implementation and evaluation, and one for related work.}
